---
title: "Bayesian_Network_Model"
format: html
editor: visual
---

## Packages

```{r}
#| echo: true
#| results: 'hide'

library(here)
library(igraph)
library(ggraph)
library(tidyr)
library(bnlearn)
library(purrr)
library(tinytex)
library(coda)
library(taxize)
library(stringr)
library(dplyr)
library(janitor)
```

## Functions

```{r}


create_resource_list <- function(links_df, nodes) {
  resource_list <- lapply(nodes, function(nm) {
    unique(links_df$from[links_df$to == nm])
  })
  names(resource_list) <- nodes
  return(resource_list)
}


bayesian_model <- function(
  n_sim,
  data,
  resource_list,
  id_col,
  alpha_a,
  alpha_b,
  max_alpha = 0.75
) {

  taxa <- names(resource_list)

  base_ext <- data$extinction_p
  names(base_ext) <- as.character(data[[id_col]])

  # Ensure ordering matches resource list
  base_ext <- base_ext[taxa]

  assign_alpha_species_bayesian <- function() {

    alpha_vec <- sapply(taxa, function(child) {

      parents <- resource_list[[child]]

      if (is.null(parents) || length(parents) == 0) {
        return(0)
      }

      rbeta(1, alpha_a, alpha_b) * max_alpha
    })

    names(alpha_vec) <- taxa

    alpha_vec
  }

  results <- matrix(
    0,
    nrow = n_sim,
    ncol = length(taxa)
  )

  colnames(results) <- taxa

  for (sim in seq_len(n_sim)) {

    alpha <- assign_alpha_species_bayesian()

    # Primary extinctions
    state <- rbinom(
      n = length(taxa),
      size = 1,
      prob = base_ext
    )

    names(state) <- taxa

    repeat {

      new_state <- state

      for (child in taxa) {

        if (state[child] == 1)
          next

        parents <- resource_list[[child]]

        if (is.null(parents))
          next

        parents <- as.character(parents)

        parents <- parents[parents %in% taxa]

        if (length(parents) == 0)
          next

        parent_state <- state[parents]

        if (sum(parent_state) == 0)
          next

        # Resource importance weights
        resource_weights <- runif(n = length(parents),
                                  min = 0,
                                  max = 1)

        # Weighted proportion of resources lost
        lost_weight <- sum(
          resource_weights * parent_state)

        total_weight <- sum(resource_weights)

        prop_resource_loss <- lost_weight / total_weight

        
    p_from_parents <- prop_resource_loss ^ (1 / alpha[child])


        p_from_parents <- min(
          max(p_from_parents, 0),
          1
        )

        if (runif(1) < p_from_parents) {

          new_state[child] <- 1

        }
      }

      # Stop when no new extinctions occur
      if (all(new_state == state))
        break

      state <- new_state
    }

    results[sim, ] <- state
  }

  results
}



posterior_summary <- function(samples_matrix) {
  mean_vals <- apply(samples_matrix, 2, mean)
  sd_vals <- apply(samples_matrix, 2, sd)
  lower_95 <- apply(samples_matrix, 2, function(x) quantile(x, 0.025))
  upper_95 <- apply(samples_matrix, 2, function(x) quantile(x, 0.975))
  
  res <- rbind(
    mean = mean_vals,
    sd = sd_vals,
    lower_95 = lower_95,
    upper_95 = upper_95
  )
  
  return(res)
}



run_bayesian_model <- function(
  resource_lists,
  data,
  n_sim,
  id_col,
  alpha_a,
  alpha_b
) {
  
  all_results <- lapply(names(resource_lists), function(web_name) {
    
    cat("Running Bayesian extinction for web:", web_name, "\n")
    
    res_list <- resource_lists[[web_name]]
    taxa <- names(res_list)
    
    # Subset node data for this network
    data <- data[data[[id_col]] %in% taxa, ]
    
    # Run Bayesian model 
    res_matrix <- bayesian_model(
      n_sim = n_sim,
      data = data,
      resource_list = res_list,
      id_col = id_col,
      alpha_a = alpha_a,
      alpha_b = alpha_b
    )
    
    # Compute posterior summaries
    posterior <- posterior_summary(res_matrix)
    
    # Convert to data frame
    df <- data.frame(
      NodeID = colnames(res_matrix),
      extinction_mean = posterior["mean", ],
      extinction_sd = posterior["sd", ],
      extinction_lower95 = posterior["lower_95", ],
      extinction_upper95 = posterior["upper_95", ],
      web = web_name
    )
    
    df
  })
  
  do.call(rbind, all_results)
}



```

# Quick Pond Food Web

## Data Import & Cleaning

```{r}

qp_link_data <- read.csv("Data/Quick_Pond_Data/Quick_Pond_Links.csv")
qp_node_data <- read.csv("Data/Quick_Pond_Data/Quick_Pond_Nodes.csv")

qp_link_data_cut <- qp_link_data


retained_nodes <- c(qp_link_data_cut$ConsumerNodeID, qp_link_data_cut$ResourceNodeID)
retained_nodes <- unique(retained_nodes)
qp_node_data_cut <- qp_node_data %>% dplyr::filter(NodeID %in% retained_nodes)
nodes <- unique(qp_node_data$NodeID)
qp_nodes_cut <- as.character(c(qp_link_data_cut$ConsumerNodeID, 
                               qp_link_data_cut$ResourceNodeID))
qp_nodes_cut <- unique(qp_nodes_cut)

qp_species_cut <- unique(as.character(c(qp_link_data_cut$ConsumerSpeciesID, qp_link_data_cut$ResourceSpeciesID)))


qp_links <- data.frame(from = qp_link_data_cut$ResourceNodeID,
                    to = qp_link_data_cut$ConsumerNodeID)

```

## Network Sorting

```{r}


qp_parasite_species <- qp_node_data_cut$SpeciesID[72:113]
qp_parasite_nodes <- qp_node_data_cut$NodeID[72:113]


qp_links_aggregated <- data.frame(from = qp_link_data_cut$ResourceSpeciesID, 
                               to = qp_link_data_cut$ConsumerSpeciesID)

qp_links_aggregated_wo_parasites <- qp_links_aggregated %>% 
  dplyr::filter(!(from %in% qp_parasite_species | to %in% qp_parasite_species))

qp_links_life_stage <- data.frame(from = qp_link_data_cut$ResourceNodeID,
                               to = qp_link_data_cut$ConsumerNodeID)

qp_links_life_stage_wo_parasites <- qp_links_life_stage %>% 
   dplyr::filter(!(from %in% qp_parasite_nodes | to %in% qp_parasite_nodes))


all_qp_food_webs <- list(
  aggregated              = qp_links_aggregated,
  aggregated_wo_parasites  = qp_links_aggregated_wo_parasites,
  life_stage               = qp_links_life_stage,
  life_stage_wo_parasites  = qp_links_life_stage_wo_parasites
)

```

## **Network Construction**

**Constructing Network & Ensuring Graph is Non-Cyclical**

```{r}

network_list <- lapply(all_qp_food_webs, function(df) {
  graph_from_data_frame(df, directed = TRUE)
})

network_list <- lapply(network_list, function(simplify) {
  igraph::simplify(simplify, remove.multiple = TRUE, remove.loops = TRUE)
})

dag_check <- lapply(network_list, function(dag_check) {
  igraph::is_dag(dag_check)
})

network_list_dag <- lapply(network_list, make_acyclic)

dag_check <- lapply(network_list_dag, function(dag_check) {
  igraph::is_dag(dag_check)
})

```

## **Assigning Baseline Extinction Probabilities**

Assigned by % threatened on the IUCN redlist.

```{r}

extinction_p <- c("Insecta" = 0.16, 
"Aves" = 0.12,
"Amphibia" = 0.41,
"Mammalia" = 0.27,
"Reptilia" = 0.21)

extinction_p[c("Bacteria", "Protista", "Fungi", "Zooplankton", 
             "Meiofauna", "Detritus", "Opalinea", "Ciliatea", "Gregarinia", 
             "Chytridiomycetes")] <- 0.10

extinction_p[c("Arachnida", "Gastropoda", "Plantae", "Secernentea", "Trematoda", "Palaeacanthocephala", "Hirudinea", "Malacostraca", "Maxillopoda", 
"Branchiopoda", "Oligochaeta", "Plant", "Secernentea", "Nematoda ")] <- 0.20


qp_node_data_cut$key <- paste(qp_node_data_cut$Class)
qp_node_data_cut$key <- ifelse(trimws(qp_node_data_cut$key) == "",
                            NA, qp_node_data_cut$key)
qp_node_data_cut$key <- ifelse(is.na(qp_node_data_cut$key),
                            qp_node_data_cut$OrganismalGroup, qp_node_data_cut$key)

qp_node_data_cut$extinction_p <- 
  extinction_p[match(qp_node_data_cut$key, names(extinction_p))]

```

## **Resource List**

```{r}


nodes_aggregated <- unique(c(all_qp_food_webs$aggregated$from,
                             all_qp_food_webs$aggregated$to))

nodes_aggregated_wo_parasites <- unique(c(all_qp_food_webs$aggregated_wo_parasites$from,
                                          all_qp_food_webs$aggregated_wo_parasites$to))

nodes_life_stage <- unique(c(all_qp_food_webs$life_stage$from,
                             all_qp_food_webs$life_stage$to))

nodes_life_stage_wo_parasites <- unique(c(all_qp_food_webs$life_stage_wo_parasites$from,
                                          all_qp_food_webs$life_stage_wo_parasites$to))


resource_lists <- list(
  aggregated = create_resource_list(all_qp_food_webs$aggregated, nodes_aggregated),
  aggregated_wo_parasites = create_resource_list(all_qp_food_webs$aggregated_wo_parasites, nodes_aggregated_wo_parasites),
  life_stage = create_resource_list(all_qp_food_webs$life_stage, nodes_life_stage),
  life_stage_wo_parasites = create_resource_list(all_qp_food_webs$life_stage_wo_parasites, nodes_life_stage_wo_parasites)
)

```

## Monte Carlo Simulation

```{r}

qp_ls_results <- run_bayesian_model(
  resource_list = resource_lists[c("life_stage", "life_stage_wo_parasites")],
  data = qp_node_data_cut,
  n_sim = 100000,
  id_col = "NodeID",
  alpha_a = 1,
  alpha_b = 1
)

qp_agg_results <- run_bayesian_model(
  resource_lists = resource_lists[c("aggregated", "aggregated_wo_parasites")],
  data = qp_node_data_cut,
  n_sim = 100000,
  id_col = "SpeciesID",
  alpha_a = 1,
  alpha_b = 1
)



```

# **Carpinteria Food Web**

## Data Import & Cleaning

```{r}


all_gw_food_webs <- readRDS("GATEWAy_Analysis/gateway_food_web_list.rds")
carp_fw <- as.data.frame(all_gw_food_webs["Carpinteria"])

carpinteria_web <- carp_fw %>%
  dplyr::select(any_of(c("Carpinteria.interaction.type", "Carpinteria.con.taxonomy",
                         "Carpinteria.con.taxonomy.level", "Carpinteria.con.common",
                         "Carpinteria.con.lifestage", "Carpinteria.res.taxonomy", 
                         "Carpinteria.res.taxonomy.level", "Carpinteria.res.common", 
                         "Carpinteria.res.lifestage", "Carpinteria.con_taxon_id", 
                         "Carpinteria.res_taxon_id")))


carp_nodes <- bind_rows(

  carpinteria_web %>%
    transmute(
      SpeciesID = Carpinteria.con_taxon_id,
      Taxonomy = Carpinteria.con.taxonomy,
      Common_Name = Carpinteria.con.common,
      Stage = Carpinteria.con.lifestage,
      Interaction = Carpinteria.interaction.type
    ),

  carpinteria_web %>%
    transmute(
      SpeciesID = NA,
      Taxonomy = Carpinteria.res.taxonomy,
      Common_Name = Carpinteria.res.common,
      Stage = Carpinteria.res.lifestage,
      Interaction = Carpinteria.interaction.type
    )

) %>%
  distinct()


carp_nodes$Interaction[is.na(carp_nodes$Interaction)] <- "parasite_consumption"
 

carp_taxon_lookup <- carp_nodes %>%
  filter(!is.na(SpeciesID)) %>%
  distinct(Taxonomy, SpeciesID)

carp_nodes <- carp_nodes %>%
  dplyr::select(-SpeciesID) %>%
  left_join(carp_taxon_lookup, by = "Taxonomy")

max_id <- max(carp_nodes$SpeciesID, na.rm = TRUE)

carp_missing_taxa <- carp_nodes %>%
  filter(is.na(SpeciesID)) %>%
  distinct(Taxonomy) %>%
  mutate(SpeciesID = max_id + row_number())

carp_nodes <- carp_nodes %>%
  dplyr::select(-SpeciesID) %>%
  left_join(
    bind_rows(carp_taxon_lookup, carp_missing_taxa),
    by = "Taxonomy"
  )

carp_nodes <- carp_nodes %>%
  distinct(SpeciesID, Taxonomy, Common_Name, Stage, Interaction) %>%
  arrange(SpeciesID, Stage) %>%
  group_by(SpeciesID) %>%
  mutate(StageID = row_number()) %>%
  ungroup()

carp_nodes <- carp_nodes %>%
  mutate(
    NodeID = row_number(),
    Species.StageID = paste0(SpeciesID, ".", StageID)
  ) %>%
  dplyr::select(
    Taxonomy,
    NodeID,
    SpeciesID,
    StageID,
    Stage,
    Species.StageID,
    Common_Name,
    Interaction
  )


carp_node_lookup <- carp_nodes %>%
  dplyr::select(
    NodeID,
    SpeciesID, 
    StageID,
    Taxonomy,
    Stage, 
    Interaction
  )


carp_edges <- carpinteria_web %>%
  left_join(
    carp_node_lookup %>%
      rename(
        ConsumerNodeID = NodeID,
        ConsumerSpeciesID = SpeciesID,
        ConsumerStageID = StageID
      ),
    by = c(
      "Carpinteria.con.taxonomy" = "Taxonomy",
      "Carpinteria.con.lifestage" = "Stage"
    )
  )

carp_edges <- carp_edges %>%
  left_join(
    carp_node_lookup %>%
      rename(
        ResourceNodeID = NodeID,
        ResourceSpeciesID = SpeciesID,
        ResourceStageID = StageID,
      ),
    by = c(
      "Carpinteria.res.taxonomy" = "Taxonomy",
      "Carpinteria.res.lifestage" = "Stage"
    )
  )

carp_edges_final <- carp_edges %>%
  dplyr::select(
    ConsumerNodeID,
    ConsumerSpeciesID,
    ResourceNodeID,
    ResourceSpeciesID,
    InteractionType = Carpinteria.interaction.type
  ) 

carp_edges_final$InteractionType[is.na(carp_edges_final$InteractionType)] <- "parasite_consumption"



carp_taxonomy_complete <- read.csv("Data_Outputs/carp_web_taxonomy.csv")

carp_nodes <- carp_nodes %>%
  mutate(
    Genus = word(Taxonomy, 1),
  )
 

carp_nodes_full <- carp_nodes %>%
  left_join(
    carp_taxonomy_complete,
    by = c("Genus" = "genus")
  ) %>%
  distinct()
 

```

```{r}

extinction_p <- c("Insecta" = 0.16, "Aves" = 0.12, "Amphibia" = 0.41, "Mammalia" = 0.27, "Reptilia" = 0.21)

extinction_p[c("Bacteria", "Protista", "Fungi", "Zooplankton", "Meiofauna", "Detritus", "Opalinea", "Ciliatea", "Gregarinia", "Chytridiomycetes", "Virus", "Apicomplexa", "Picornavirus")] <- 0.10

extinction_p[c("Arachnida", "Gastropoda", "Plantae", "Secernentea", "Trematoda", "Palaeacanthocephala", "Hirudinea", "Malacostraca", "Maxillopoda", 
"Branchiopoda", "Oligochaeta", "Plantae", "Secernentea", "Nematoda", "Polychaeta", "Teleostei", "Chondrichthyes", "Polychaeta", "Neonemertea", "Palaeonemertea","Gastropoda", "Cnidaria", "Mollusca", "Aconoidasida", "Acentrosomata", "Chromadorea", "Branchiura", "Oligohymenophorea", "Myxozoa", "Bivalvia", "Digenea", "Magnoliopsida", "Liliopsida", "Tracheophyta")] <- 0.20

carp_nodes_full[carp_nodes_full == "NA"] <- NA

carp_nodes_full$key <- dplyr::coalesce(
  carp_nodes_full$class,
  carp_nodes_full$phylum,
  carp_nodes_full$kingdom,
  carp_nodes_full$Genus
)

carp_nodes_full$extinction_p <- 
  extinction_p[match(carp_nodes_full$key, names(extinction_p))]



```

```{r}


carp_parasitic_species_ids <- carp_nodes_full %>%
  filter(
    phylum %in% c("Platyhelminthes", "Nematoda", "Nemertea") |
    key %in% c("Apicomplexa", "Picornavirus")
) %>%
  pull(SpeciesID) %>%
  unique()


carp_parasitic_node_ids <- carp_nodes_full %>%
  filter(
    phylum %in% c("Platyhelminthes", "Nematoda", "Nemertea") |
    key %in% c("Apicomplexa", "Picornavirus")
) %>%
  pull(NodeID) %>%
  unique()


```

```{r}

carpinteria_aggregated <- data.frame(from = carp_edges_final$ResourceSpeciesID, 
                                     to = carp_edges_final$ConsumerSpeciesID) %>% unique()

carpinteria_aggregated_wo_parasites <- data.frame(
  from = carp_edges_final$ResourceSpeciesID,
  to   = carp_edges_final$ConsumerSpeciesID
) %>%
  filter(
    !(from %in% carp_parasitic_species_ids |
      to %in% carp_parasitic_species_ids)
  ) %>%
  distinct()
 
carpinteria_disaggregated <- data.frame(from = carp_edges_final$ResourceNodeID, 
                                         to = carp_edges_final$ConsumerNodeID) %>% unique()

carpinteria_disaggregated_wo_parasites <- data.frame(
  from = carp_edges_final$ResourceNodeID,
  to = carp_edges_final$ConsumerNodeID) %>% 
  filter(
    !(from %in% carp_parasitic_node_ids |
      to %in% carp_parasitic_node_ids)
  ) %>%
  distinct()


all_carpinteria_edge_lists <- list(
  carpinteria_aggregated = carpinteria_aggregated, 
  carpinteria_aggregated_wo_parasites = carpinteria_aggregated_wo_parasites, 
  carpinteria_disaggregated = carpinteria_disaggregated, 
  carpinteria_disaggregated_wo_parasites = carpinteria_disaggregated_wo_parasites
)


all_carpinteria_networks <- lapply(all_carpinteria_edge_lists, function(graph) {
  graph_from_data_frame(graph)
})

all_carpinteria_networks <- lapply(all_carpinteria_networks, function(simplify) {
  igraph::simplify(simplify, remove.multiple = TRUE, remove.loops = TRUE)
})

dag_check <- lapply(all_carpinteria_networks, function(dag_check) {
  igraph::is_dag(dag_check)
})


all_carpinteria_networks_dag <- lapply(all_carpinteria_networks, make_acyclic)

dag_check <- lapply(all_carpinteria_networks_dag, function(dag_check) {
  igraph::is_dag(dag_check)
})


all_carpinteria_matrices <- all_carpinteria_networks %>%
  map(~ igraph::as_adjacency_matrix(.x, sparse = FALSE))


```

## **Assigning Baseline Extinction Probabilities**

Assigned by % threatened on the IUCN redlist.

```{r}



hierarchy_full <- function(taxonomy, db = "itis") {
  uids <- get_tsn(taxonomy, messages = FALSE)
  classifications <- taxize::classification(tsns, db = db)
  taxonomy_df <- do.call(rbind, lapply(names(classifications), function(taxonomy) {
    cl <- classifications[[taxonomy]]
    if (!is.null(cl)) {
      cl$taxonomy <- taxonomy
      return(cl)
    } else {
      return(NULL)
    }
  }))
  
  return(taxonomy_df)
}

##carp_taxa <- c(carp_nodes$Genus)

##taxonomy <- taxize::classification(as.character(carp_taxa), db = "itis")


taxonomy_df <- purrr::imap_dfr(taxonomy, ~{
  .x %>%
    dplyr::filter(rank) %>%
    dplyr::mutate(taxon = .y)
}) %>%
  tidyr::pivot_wider(
    id_cols = taxon,
    names_from = rank,
    values_from = name
  ) %>%
  dplyr::select(taxon, all_of(ranks_keep))


taxonomy_df <- taxonomy_df %>%
  dplyr::select(
    taxon,
    any_of(c(
      "kingdom",
      "phylum",
      "class",
      "order",
      "family",
      "genus"
    ))
  )

```

## **Resource List**

```{r}


carp_nodes_agg <- carp_nodes_full %>%
  distinct(SpeciesID, .keep_all = TRUE)

carp_nodes_agg_wo_parasites <- carp_nodes_agg %>% filter(Interaction %in% c("predacious", "parasite_consumption")) %>% filter(!phylum %in% c("Platyhelminthes", "Nemertea"))  %>% 
                                                                                                                                  filter(!key %in% c("Picornavirus", "Apicomplexa"))


carp_nodes_full_wo_parasites <- carp_nodes_full %>% filter(Interaction %in% c("predacious", "parasite_consumption")) %>% filter(!phylum %in% c("Platyhelminthes", "Nemertea")) %>% 
                                                                                                                                  filter(!key %in% c("Picornavirus", "Apicomplexa"))


carp_resource_lists <- list(
  aggregated = create_resource_list(carpinteria_aggregated, 
                                    node_ids = carp_nodes_agg$SpeciesID), 
  aggregated_wo_parasites = create_resource_list(carpinteria_aggregated_wo_parasites, 
                                          node_ids = carp_nodes_agg_wo_parasites$SpeciesID),
  life_stage = create_resource_list(carpinteria_disaggregated, 
                                    node_ids = carp_nodes_full$NodeID),
  life_stage_wo_parasites = create_resource_list(carpinteria_disaggregated_wo_parasites, node_ids = carp_nodes_full_wo_parasites$NodeID)
)


```

## **Monte Carlo Simulation**

```{r}

carp_agg_results <- run_bayesian_model(
  resource_list = carp_resource_lists[c("aggregated")],
  data = carp_nodes_agg,
  n_sim = 10000,
  id_col = "SpeciesID",
  alpha_a = 1,
  alpha_b = 1
)


carp_agg_wo_parasites_results <- run_bayesian_model(
  resource_list = carp_resource_lists[c("aggregated_wo_parasites")],
  data = carp_nodes_agg_wo_parasites,
  n_sim = 100000,
  id_col = "SpeciesID",
  alpha_a = 1,
  alpha_b = 1
)

carp_ls_results <- run_bayesian_model(
  resource_list = carp_resource_lists[c("life_stage")],
  data = carp_nodes_full,
  n_sim = 100000,
  id_col = "NodeID",
  alpha_a = 1,
  alpha_b = 1
)

carp_ls_wo_parasites_results <- run_bayesian_model(
  resource_list = carp_resource_lists[c("life_stage_wo_parasites")],
  data = carp_nodes_full,
  n_sim = 100000,
  id_col = "NodeID",
  alpha_a = 1,
  alpha_b = 1
)




```

# Ythan Estuary Food Web

## Data Import & Cleaning

```{r}


ythan_nodes <- read.csv("Data/Dunne_2013/Dunne_2013_Species_Lists.csv") %>% 
  dplyr::select(25:27)

ythan_nodes <- row_to_names(ythan_nodes, row_number=1) %>%
   filter(if_any(everything(), ~ !is.na(.) & . != ""))

ythan_links_all <- read.csv("Data/Dunne_2013/Ythan_Estuary_Links.csv")

ythan_parasites_removed <- ythan_links_all %>% dplyr::select(1:2) %>%
   filter(if_any(everything(), ~ !is.na(.) & . != ""))
ythan_parasites_included <- ythan_links_all %>% dplyr::select(4:5) %>% 
   filter(if_any(everything(), ~ !is.na(.) & . != ""))
ythan_parasites_concamitant <- ythan_links_all %>% dplyr::select(7:8) %>% 
   filter(if_any(everything(), ~ !is.na(.) & . != ""))

colnames(ythan_parasites_removed) <- c("to", "from")
colnames(ythan_parasites_included) <- c("to", "from")
colnames(ythan_parasites_concamitant) <- c("to", "from")
```

```{r}

ythan_nodes <- ythan_nodes %>%
  mutate(
    Genus = word(Name, 1),
  )
```

```{r}



hierarchy_full <- function(taxonomy, db = "itis") {
  uids <- get_tsn(taxonomy, messages = FALSE)
  classifications <- taxize::classification(tsns, db = db)
  taxonomy_df <- do.call(rbind, lapply(names(classifications), function(taxonomy) {
    cl <- classifications[[taxonomy]]
    if (!is.null(cl)) {
      cl$taxonomy <- taxonomy
      return(cl)
    } else {
      return(NULL)
    }
  }))
  
  return(taxonomy_df)
}


ythan_taxa <- c(ythan_nodes$Genus)

ythan_taxonomy <- taxize::classification(as.character(ythan_taxa), db = "itis")



ythan_taxonomy_df <- imap_dfr(
  ythan_taxonomy[!sapply(ythan_taxonomy, function(x) all(is.na(x)))],
  ~ .x %>%
      dplyr::select(rank, name) %>%
      pivot_wider(
        names_from = rank,
        values_from = name
      ) %>%
      mutate(taxon = .y, .before = 1)
) %>%
  dplyr::select(c("kingdom", "phylum", "class", "order", "family", "genus"))

ythan_nodes <- left_join(ythan_nodes,ythan_taxonomy_df, by = c("Genus" = "genus"))

ythan_nodes_complete <- read.csv("Data_Outputs/ythan_nodes_complete.csv")


extinction_p <- c("Insecta" = 0.16, "Aves" = 0.12, "Amphibia" = 0.41, "Mammalia" = 0.27, "Reptilia" = 0.21)

extinction_p[c("Bacteria", "Protista", "Fungi", "Zooplankton", "Meiofauna", "Detritus", "Opalinea", "Ciliatea", "Gregarinia", "Chytridiomycetes", "Virus", "Apicomplexa", "Picornavirus", "Apicomplexan")] <- 0.10

extinction_p[c("Arachnida", "Gastropoda", "Plantae", "Secernentea", "Trematoda", "Palaeacanthocephala", "Hirudinea", "Malacostraca", "Maxillopoda", 
"Branchiopoda", "Oligochaeta", "Plantae", "Secernentea", "Nematoda", "Polychaeta", "Teleostei", "Chondrichthyes", "Polychaeta", "Neonemertea", "Palaeonemertea","Gastropoda", "Cnidaria", "Mollusca", "Aconoidasida", "Acentrosomata", "Chromadorea", "Branchiura", "Oligohymenophorea", "Myxozoa", "Bivalvia", "Digenea", "Magnoliopsida", "Liliopsida", "Tracheophyta", "Platyhelminthes", "Teleostei", "Arthropoda", "Clitellata", "Acanthocephala", "Thecostraca", "Annelida", 
"Echinodermata", "Nermertea", "Phoronid", "Bivalve", "Crustacea", "Actinopterygii", "Isopoda", "Ciliophora")] <- 0.20

extinction_p["POM"] <- 0

ythan_nodes_complete$extinction_p <- 
  extinction_p[match(ythan_nodes_complete$Key, names(extinction_p))]
  
  
ythan_nodes_complete <- ythan_nodes_complete %>%
  distinct(ID, .keep_all = TRUE)

ythan_nodes_complete_wo_parasites <- ythan_nodes_complete %>% filter(!ID %in% c(1:42, 108, 110, 111, 113, 125, 126, 130, 132)) %>%
   distinct(ID, .keep_all = TRUE)

```

## **Resource List**

```{r}



ythan_resource_lists <- list(
  without_parasites = create_resource_list(ythan_parasites_removed, 
                                           nodes = ythan_nodes_complete_wo_parasites$ID),
  with_parasites_concamitant = create_resource_list(ythan_parasites_concamitant, 
                                                    nodes = ythan_nodes_complete$ID)
  )



```

## **Monte Carlo Simulation**

```{r}


ythan_without_parasites_results <- run_bayesian_model(
  resource_list = ythan_resource_lists[c("without_parasites")],
  data = ythan_nodes_complete_wo_parasites,
  n_sim = 100000,
  id_col = "ID",
  alpha_a = 1,
  alpha_b = 1
)

ythan_with_parasites_concam_results <- run_bayesian_model(
  resource_list = ythan_resource_lists[c("with_parasites_concamitant")],
  data = ythan_nodes_complete,
  n_sim = 100000,
  id_col = "ID",
  alpha_a = 1,
  alpha_b = 1
)


```

# Punta Banda Food Web

## Data Import & Cleaning

```{r}


banda_nodes <- read.csv("Data/Dunne_2013/Dunne_2013_Species_Lists.csv") %>% 
  dplyr::select(9:11)
banda_nodes <- row_to_names(banda_nodes, row_number=1) %>%
   filter(if_any(everything(), ~ !is.na(.) & . != ""))


banda_links_all <- read.csv("Data/Dunne_2013/Bahia_Falsa_Web.csv")


banda_parasites_removed <- banda_links_all %>% dplyr::select(1:2) %>%
   filter(if_any(everything(), ~ !is.na(.) & . != ""))
banda_parasites_included <- banda_links_all %>% dplyr::select(4:5) %>% 
   filter(if_any(everything(), ~ !is.na(.) & . != ""))
banda_parasites_concamitant <- banda_links_all %>% dplyr::select(7:8) %>% 
   filter(if_any(everything(), ~ !is.na(.) & . != ""))

colnames(banda_parasites_removed) <- c("to", "from")
colnames(banda_parasites_included) <- c("to", "from")
colnames(banda_parasites_concamitant) <- c("to", "from")






```

```{r}


banda_nodes_complete <- read.csv("Data_Outputs/banda_nodes_complete.csv")



banda_nodes_complete$extinction_p <- extinction_p[match(banda_nodes_complete$key, names(extinction_p))]

banda_nodes_complete_wo_parasites <- banda_nodes_complete %>% filter(!ID %in% c(15:29, 25, 139:147, 149:209, 214))

```

## **Resource List**

```{r}


banda_resource_lists <- list(
  without_parasites = create_resource_list(banda_parasites_removed, 
                                           nodes = banda_nodes_complete_wo_parasites$ID),
  with_parasites_concamitant = create_resource_list(banda_parasites_concamitant, 
                                                    nodes = banda_nodes_complete$ID)
  )

```

## **Monte Carlo Simulation**

```{r}



banda_without_parasites_results <- run_bayesian_model(
  resource_list = banda_resource_lists[c("without_parasites")],
  data = banda_nodes_complete,
  n_sim = 100000,
  id_col = "ID",
  alpha_a = 1,
  alpha_b = 1
)

banda_with_parasites_concam_results <- run_bayesian_model(
  resource_list = banda_resource_lists[c("with_parasites_concamitant")],
  data = banda_nodes_complete,
  n_sim = 100000,
  id_col = "ID",
  alpha_a = 1,
  alpha_b = 1
)

```
