---
title: "Parasite_Food_Web_Topological_Analysis"
author: "Finlay Ryder"
format: html
editor: visual
---

# Food Web Topological Analysis

## Packages

```{r}
library(here)
library(tidyr)
library(dplyr)
library(igraph)
library(NetIndices)
library(purrr)
library(NetworkExtinction)
library(pracma)
library(lhs)
library(vegan)
library(stringr)

```

## **Functions**

```{r}


make_square <- function(mat) {
  if (is.null(rownames(mat)))
    rownames(mat) <- paste0("R_", seq_len(nrow(mat)))
  if (is.null(colnames(mat)))
    colnames(mat) <- paste0("C_", seq_len(ncol(mat)))
  all_nodes <- c(rownames(mat), colnames(mat))
  square_mat <- matrix(
    0,
    nrow = length(all_nodes),
    ncol = length(all_nodes),
    dimnames = list(all_nodes, all_nodes)
  )
  square_mat[rownames(mat), colnames(mat)] <- mat
  square_mat
}


compute_robustness_auc_most <- function(mat) {
  res_most <- SimulateExtinctions(
    Network = mat,
    Method = "Mostconnected",
    NetworkType = "Trophic")
  x   <- res_most$sims$NumExt / res_most$sims$S[1]
  y   <- res_most$sims$S      / res_most$sims$S[1]
  AUC_most <- pracma::trapz(x, y)
  return(
    list(
      Most  = res_most,
      AUC_most  = AUC_most
    )
  )
}


compute_robustness_auc_random <- function(mat) {
  set.seed(100)
  res_ran <- RandomExtinctions(
    Network = mat,
    nsim = 10,
    plot = FALSE  
  )
  NumExt <- res_ran$sims$NumExt
  AccSecExt <- res_ran$sims$AccSecExt_mean
  TotalExt <- NumExt + AccSecExt
  TotalExt <- cummax(TotalExt)
  S0 <- max(TotalExt)
  x <- NumExt / S0
  y <- (S0 - TotalExt) / S0
  AUC_random <- pracma::trapz(x, y)
  return(list(
    random = res_ran,
    AUC_random = AUC_random
  ))
}


reindex_web <- function(web) {
  rownames(web) <- colnames(web) <- seq_len(nrow(web))
  web
}


generate_null <- function(web){

  S <- nrow(web)
  L <- sum(web)

  ##random trophic ordering
  ord <- sample.int(S)

  ##finding all possible links
  possible_links <- which(
    upper.tri(matrix(0, S, S)),
    arr.ind = TRUE
  )

  ##possible predator-prey links given trophic ordering, for L links
  chosen <- possible_links[
    sample(nrow(possible_links), L),
    ,
    drop = FALSE
  ]
  ##constructs null webs based upon chosen link options
  null_web <- matrix(0, S, S)

  for(k in seq_len(nrow(chosen))){

    prey <- ord[chosen[k, 1]]
    predator <- ord[chosen[k, 2]]

    null_web[predator, prey] <- 1
  }

  return(null_web)
}

igraph_to_network <- function(g) {
  intergraph::asNetwork(g)
}

##runs calculation of common food web metrics for input networks
foodweb_metrics <- function(web){
  
   g <- graph_from_adjacency_matrix(
    web,
    mode = "directed"
  )

  troph <- TrophInd(web)

  data.frame(

    Species = nrow(web),
    Links = sum(web),
    Connectance = sum(web)/(nrow(web)^2),
    
    Nestedness = nestednodf(web)$statistic["NODF"],

    Mean_degree = mean(igraph::degree(g)),
    SD_degree = sd(igraph::degree(g)),
    Max_degree = max(igraph::degree(g)),

    Clustering = transitivity(g, type = "global"),

    Modularity = modularity(
      cluster_louvain(as.undirected(g))
    ),

    Mean_TL = mean(troph$TL),
    Max_TL = max(troph$TL),

    Mean_Omnivory = mean(troph$OI, na.rm = TRUE),

    Mean_Path_Length = mean_distance(
      g,
      directed = TRUE
    )
  )
}


network_metrics <- c(
  "Species",
  "Links",
  "Connectance", 
  "Nestedness", 
  "Mean_degree",
  "SD_degree",
  "Max_degree", 
  "Clustering", 
  "Modularity", 
  "Mean_TL", 
  "Max_TL", 
  "Mean_Omnivory", 
  "Mean_Path_Length"
)


##Selects 10 networks from all generated null networks which maximise span across metric space, for metrics calculated for each null network using `foodweb_metrics()` function. 
maximin_select <- function(df, n_select = 10) {

  set.seed(100)

  ##drops metrics with almost 0 variance (i.e., those metrics deliberately costrained), and standardised variables so distance of one isn't more important, and the model, than other because it has a larger scale. 
  X <- df %>%
    dplyr::select(where(is.numeric)) %>%
    dplyr::select(where(~ sd(.x, na.rm = TRUE) > 1e-10)) %>%
    scale()

  X <- as.matrix(X)
##selects first row randomly. 
  selected <- sample(seq_len(nrow(X)), 1)
##keeps going until 10 networks are selected
  while(length(selected) < n_select) {

    remaining <- setdiff(seq_len(nrow(X)), selected)
##finds minimum distances between all, and then finds the maximum distance of all minima calculated
    min_dist <- sapply(remaining, function(i) {

      min(
        rowSums(
          (
            X[selected, , drop = FALSE] -
            matrix(
              X[i, ],
              nrow = length(selected),
              ncol = ncol(X),
              byrow = TRUE
            )
          )^2
        )
      )

    })

    selected <- c(
      selected,
      remaining[which.max(min_dist)]
    )
  }

  selected
}


```

## **Loading & Cleaning Food Webs**

```{r}

all_gw_food_webs <- readRDS("GATEWAy_Analysis/gateway_food_web_list.rds")
carp_fw <- as.data.frame(all_gw_food_webs["Carpinteria"])

carpinteria_web <- carp_fw %>%
  dplyr::select(any_of(c("Carpinteria.interaction.type", "Carpinteria.con.taxonomy",
                         "Carpinteria.con.taxonomy.level", "Carpinteria.con.common",
                         "Carpinteria.con.lifestage", "Carpinteria.res.taxonomy", 
                         "Carpinteria.res.taxonomy.level", "Carpinteria.res.common", 
                         "Carpinteria.res.lifestage", "Carpinteria.con_taxon_id", 
                         "Carpinteria.res_taxon_id")))


quick_pond_adj_matrices <- readRDS("Data/Quick_Pond_Data/quick_pond_adj_matrices.rds")
names(quick_pond_adj_matrices) <- c("qp_agg", "qp_agg_wo_parasites", 
                                    "qp_lifestage", "qp_lifestages_wo_parasites")





```

## Quick Pond Food Web

### **Partitioning Webs + Make Webs Acyclical**

```{r}


all_qp_networks <- lapply(quick_pond_adj_matrices, function(graph) {
  graph_from_adjacency_matrix(graph)
})


all_qp_networks <- lapply(all_qp_networks, function(simplify) {
  igraph::simplify(simplify, remove.multiple = TRUE, remove.loops = TRUE)
})

dag_check <- lapply(all_qp_networks, function(dag_check) {
  igraph::is_dag(dag_check)
})


dag_check <- lapply(all_qp_networks_dag, function(dag_check) {
  igraph::is_dag(dag_check)
})


all_qp_matrices <- all_qp_networks_dag %>%
  map(~ igraph::as_adjacency_matrix(.x, sparse = FALSE))


all_qp_matrices_cleaned <- map(all_qp_matrices, make_square)



```

Calculates network metrics.

```{r}

qp_metrics <- all_qp_matrices %>%
  map(~ foodweb_metrics(.x))

qp_ascendency <- all_qp_matrices %>%
  map(~ AscInd(.x))

```

### **Extinction Simulations**

```{r}

##ensures that networks are fully connected
sapply(all_qp_matrices_cleaned, function(mat) {
  g <- igraph::graph_from_adjacency_matrix(mat)
  c(
    nodes = vcount(g),
    edges = ecount(g),
    is_connected = igraph::is_connected(g)
  )
})

qp_robustness_results_auc_most <- all_qp_matrices_cleaned %>%
  map(compute_robustness_auc_most) 

qp_robustness_results_auc_most_df <- read.csv("Data_Outputs/Extinction_Sim/Quick_Pond/QP_MC_Sim_Results.csv")


qp_robustness_results_auc_random <- readRDS("Data_Outputs/qp_robustness_results_auc_random.rds")

qp_robustness_results_auc_random_df <- read.csv("Data_Outputs/Extinction_Sim/qp_robustness_results_auc_random.csv")


```

### **Null Model Runs**

```{r}


qp_selected_null_networks <- readRDS("Data_Outputs/Null_Models/qp_selected_networks.rds")

null_qp_robustness_results_auc_most <-
  qp_selected_null_networks %>%
  map(~ map(.x, compute_robustness_auc_most))

null_qp_robustness_results_auc_most_df <- 
  read.csv("Data_Outputs/Null_Models/null_qp_robustness_MC_results.csv")



```

## Carpinteria Food Web

### **Annotating Life Stages**

```{r}

carpinteria_stage_table <- bind_rows(carpinteria_web %>%
                           dplyr::select(taxon_id = Carpinteria.con_taxon_id, 
                                         lifestage = Carpinteria.con.lifestage),
                         
                         carpinteria_web %>%
                           dplyr::select(taxon_id = Carpinteria.res_taxon_id, 
                                         lifestage = Carpinteria.res.lifestage)) %>%
  distinct() %>%
  arrange(taxon_id, lifestage) %>%
  mutate(stage_id = row_number())


carpinteria_web <- carpinteria_web %>%
  left_join(carpinteria_stage_table, 
            by = c("Carpinteria.con_taxon_id" = "taxon_id",
           "Carpinteria.con.lifestage" = "lifestage")) %>%
  rename(con_stage_id = stage_id) %>%
             left_join(carpinteria_stage_table, 
            by = c("Carpinteria.res_taxon_id" = "taxon_id",
           "Carpinteria.res.lifestage" = "lifestage")) %>%
  rename(res_stage_id = stage_id)



  

```

### **Partitioning Webs + Make Webs Acyclical**

```{r}


carp_parasitic_species_ids <- carp_nodes_full %>%
  filter(
    phylum %in% c("Platyhelminthes", "Nematoda", "Nemertea") |
    key %in% c("Apicomplexa", "Picornavirus")
) %>%
  pull(SpeciesID) %>%
  unique()


carp_parasitic_node_ids <- carp_nodes_full %>%
  filter(
    phylum %in% c("Platyhelminthes", "Nematoda", "Nemertea") |
    key %in% c("Apicomplexa", "Picornavirus")
) %>%
  pull(NodeID) %>%
  unique()



carpinteria_wo_parasites <- carpinteria_web %>%
  filter(Carpinteria.interaction.type == "predacious")


carpinteria_aggregated <- data.frame(from = carpinteria_web$Carpinteria.res_taxon_id, 
                                     to = carpinteria_web$Carpinteria.con_taxon_id)

carpinteria_aggregated_wo_parasites <- data.frame(
  from = carpinteria_wo_parasites$Carpinteria.res_taxon_id,
  to = carpinteria_wo_parasites$Carpinteria.con_taxon_id)

carpineteria_disaggregated <- data.frame(from = carpinteria_web$res_stage_id, 
                                         to = carpinteria_web$con_stage_id)

carpinteria_disaggregated_wo_parasites <- data.frame(
  from = carpinteria_wo_parasites$res_stage_id,
  to = carpinteria_wo_parasites$con_stage_id)


all_carpinteria_edge_lists <- list(
  carpinteria_aggregated = carpinteria_aggregated, 
  carpinteria_aggregated_wo_parasites = carpinteria_aggregated_wo_parasites, 
  carpineteria_disaggregated = carpineteria_disaggregated, 
  carpinteria_disaggregated_wo_parasites = carpinteria_disaggregated_wo_parasites
)


all_carpinteria_networks <- lapply(all_carpinteria_edge_lists, function(graph) {
  graph_from_data_frame(graph)
})

all_carpinteria_networks <- lapply(all_carpinteria_networks, function(simplify) {
  igraph::simplify(simplify, remove.multiple = TRUE, remove.loops = TRUE)
})

dag_check <- lapply(all_carpinteria_networks, function(dag_check) {
  igraph::is_dag(dag_check)
})


dag_check <- lapply(all_carpinteria_networks_dag, function(dag_check) {
  igraph::is_dag(dag_check)
})


all_carpinteria_matrices <- all_carpinteria_networks %>%
  map(~ igraph::as_adjacency_matrix(.x, sparse = FALSE))


all_carpinteria_matrices_cleaned <- map(all_carpinteria_matrices, make_square)
all_carpinteria_matrices_cleaned <- 
  map(all_carpinteria_matrices_cleaned, reindex_web)


```

```{r}

all_carp_matrices <- readRDS("Data_Outputs/all_carpinteria_matrices.rds")
```

```{r}

carp_metrics <- all_carp_matrices %>% map(~ foodweb_metrics(.x))

carp_ascendency <- all_carp_matrices %>% map(~ AscInd(.x))

```

### **Extinction Simulation**

```{r}

sapply(all_carpinteria_matrices_cleaned, function(mat) {
  g <- igraph::graph_from_adjacency_matrix(mat)
  c(
    nodes = vcount(g),
    edges = ecount(g),
    is_connected = igraph::is_connected(g)
  )
})


robustness_results_auc_most <- all_carpinteria_matrices_cleaned %>%
  map(compute_robustness_auc_most) 

robustness_results_auc_random <- all_carpinteria_matrices_cleaned %>% 
  map(compute_robustness_auc_random) 

carp_robustness_results_auc_most_df <- read.csv("Data_Outputs/Extinction_Sim/Carpinteria/Carp_MC_Sim_Results.csv")





```

### **Null Model Runs**

```{r}

carp_selected_null_networks <- readRDS("Data_Outputs/Null_Models/carp_selected_networks.rds")

null_carp_robustness_results_auc_most <-
  carp_selected_null_networks %>%
  map(~ map(.x, compute_robustness_auc_most))


null_carp_robustness_most_results_df <- read.csv("Data_Outputs/Null_Models/null_carp_robustness_MC_results.csv")
```

## Ythan Estuary Food Web

```{r}


ythan_parasites_included <- readRDS("Data/Dunne_2013/ythan_parasites_concamitant.rds")
ythan_parasites_removed <- readRDS("Data/Dunne_2013/ythan_parasites_removed.rds")

ythan_parasites_included_web <- graph_from_data_frame(ythan_parasites_included, 
                                                      directed = TRUE)
ythan_parasites_removed_web <- graph_from_data_frame(ythan_parasites_removed, 
                                                      directed = TRUE)

all_ythan_webs <- list(
  ythan_parasites_included_web,
  ythan_parasites_removed_web
)

all_ythan_webs <- all_ythan_webs %>%
  purrr::map(~ igraph::simplify(
    .x,
    remove.multiple = TRUE,
    remove.loops = TRUE
  ))

all_ythan_mats <- all_ythan_webs %>%
  map(as_adjacency_matrix, sparse = FALSE)

all_ythan_mats <- all_ythan_mats %>%
  map(make_square)


ythan_metrics_parasites_included<- GenInd(ythan_parasites_included_mat)
ythan_asc_parasites_included <- AscInd(ythan_parasites_included_mat)

ythan_metrics_parasites_removed<- GenInd(ythan_parasites_removed_mat)
ythan_asc_parasites_removed <- AscInd(ythan_parasites_removed_mat)

  
  



```

### **Extinction Simulations**

```{r}


sapply(all_ythan_webs, function(mat) {
  g <- igraph::graph_from_adjacency_matrix(mat)
  c(
    nodes = vcount(g),
    edges = ecount(g),
    is_connected = igraph::is_connected(g)
  )
})



ye_robustness_results_auc_most <- all_ythan_mats %>% 
  map(compute_robustness_auc_most)

ye_robustness_results_auc_most_df <- read.csv("Data_Outputs/Extinction_Sim/Ythan/ye_MC_Sim_Results.csv")

ye_robustness_results_auc_ran <- readRDS("Data_Outputs/Extinction_Sim/Ythan/ye_robustness_results_auc_random.rds")

```

### **Null Model Runs**

```{r}


ye_selected_null_networks <- readRDS("Data_Outputs/Null_Models/ye_selected_networks.rds")

null_ye_robustness_results_auc_most <-
  ye_selected_null_networks %>%
  map(~ map(.x, compute_robustness_auc_most))

null_ye_robustness_results_auc_most_df <- read.csv("Data_Outputs/Null_Models/null_ye_robustness_MC_results.csv")

```

## Punta Banda Food Web

```{r}

banda_parasites_included <- readRDS("Data/Dunne_2013/banda_parasites_concamitant.rds")
banda_parasites_removed <- readRDS("Data/Dunne_2013/banda_parasites_removed.rds")


banda_parasites_included_web <- graph_from_data_frame(banda_parasites_included, 
                                                      directed = TRUE)
banda_parasites_removed_web <- graph_from_data_frame(banda_parasites_removed, 
                                                      directed = TRUE)

all_banda_webs <- list(
  banda_parasites_included_web, 
  banda_parasites_removed_web
)

all_banda_webs <- all_banda_webs %>%
  purrr::map(~ igraph::simplify(
    .x,
    remove.multiple = TRUE,
    remove.loops = TRUE
  ))


all_banda_mats <- all_banda_webs %>%
  map(as_adjacency_matrix, sparse = FALSE)

all_banda_mats <- all_banda_mats %>%
  map(make_square)





banda_metrics_parasites_included <- GenInd(banda_parasites_included_mat)
banda_asc_parasites_included <- AscInd(banda_parasites_included_mat)

banda_metrics_parasites_removed <- GenInd(banda_parasites_removed_mat)
banda_asc_parasites_removed <- AscInd(banda_parasites_removed_mat)

```

### **Extinction Simulations**

```{r}

pb_robustness_results_auc_most <- all_banda_mats %>% 
  map(compute_robustness_auc_most)

pb_robustness_results_auc_most_df <- read.csv("Data_Outputs/Extinction_Sim/Banda/pb_MC_Sim_Results.csv")


pb_robustness_results_auc_ran <- readRDS("Data_Outputs/Extinction_Sim/Banda/banda_robustness_results_auc_random.rds")


```

### **Null Model Runs**

```{r}

pb_selected_null_networks <- readRDS("Data_Outputs/Null_Models/pb_selected_networks.rds")

null_pb_robustness_results_auc_most <-
  pb_selected_null_networks %>%
  map(~ map(.x, compute_robustness_auc_most))
  
null_pb_robustness_results_auc_most_df <- read.csv("Data_Outputs/Null_Models/null_pb_robustness_MC_results.csv")
  
```

## All Z-Score Analysis

```{r}

qp_most_null_stats <- null_qp_robustness_results_auc_most_df %>%
  group_by(Web) %>%
  summarise(
    null_mean = mean(AUC),
    null_sd = sd(AUC),
    .groups = "drop"
  )


qp_most_z_scores <- qp_robustness_results_auc_most_df %>%
  left_join(qp_most_null_stats, by = c("Network" = "Web")) %>%
  mutate(
    z_score = (AUC - null_mean) / null_sd
  )


carp_null_stats <- null_carp_robustness_most_results_df %>%
  group_by(Web) %>%
  summarise(
    null_mean = mean(AUC),
    null_sd = sd(AUC),
    .groups = "drop"
  )

carp_z_scores <- carp_robustness_results_auc_most_df %>%
  left_join(carp_null_stats, by = c("Network" = "Web")) %>%
  mutate(
    z_score = (AUC - null_mean) / null_sd
  )


ye_null_stats <- null_ye_robustness_results_auc_most_df %>%
  group_by(Web) %>%
  summarise(
    null_mean = mean(AUC),
    null_sd = sd(AUC),
    .groups = "drop"
  )

ye_z_scores <- ye_robustness_results_auc_most_df %>%
  left_join(ye_null_stats, by = c("Network" = "Web")) %>%
  mutate(
    z_score = (AUC - null_mean) / null_sd
  )


pb_null_stats <- null_pb_robustness_results_auc_most_df %>%
  group_by(Web) %>%
  summarise(
    null_mean = mean(AUC),
    null_sd = sd(AUC),
    .groups = "drop"
  )

pb_z_scores <- pb_robustness_results_auc_most_df %>%
  left_join(pb_null_stats, by = c("Network" = "Web")) %>%
  mutate(
    z_score = (AUC - null_mean) / null_sd
  )



all_z_scores <- bind_rows(qp_most_z_scores, carp_z_scores, ye_z_scores, pb_z_scores)

all_z_scores <- all_z_scores %>%
  mutate(
    Parasites = if_else(
      str_detect(Network, "wo_parasites|without_parasites"),
      "Without parasites",
      "With parasites"
    )
  ) %>%
  
  mutate(
    FoodWeb = case_when(
      grepl("qp_agg", Network) ~ "QP (Agg)",
      grepl("qp_lifestage", Network) ~ "QP (LS)",
      grepl("carp_agg", Network) ~ "CSM (Agg)",
      grepl("carp_lifestage", Network) ~ "CSM (LS)", 
      grepl("pb_with_parasites", Network) ~ "PB",
      grepl("pb_without_parasites", Network) ~ "PB",
      grepl("ye_with_parasites", Network) ~ "YE",
      grepl("ye_without_parasites", Network) ~ "YE",
      
    )
  )



png(height = 400, width = 750)
ggplot(all_z_scores,
       aes(x = FoodWeb,
           y = z_score,
           colour = Parasites,
           group = FoodWeb)) +
  geom_line(colour = "grey50", linewidth = 0.8) +
  geom_point(size = 5) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  scale_colour_manual(
    values = c(
      "With parasites" = "red",
      "Without parasites" = "blue"
    )
  ) +
  ylim(-8, 2.5) +
  labs(y = "Z-Scores", x = "Food web", colour = NULL) +
  theme(
    axis.title = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 20, face = "bold"),
    axis.text.x = element_text(
      size = 15,
      face = "bold",
      angle = 45,
      hjust = 1
    ),
    axis.title.x = element_blank(),
    legend.position = "none"
  )

dev.off()
```
